%File: main.tex
\documentclass[letterpaper]{article} % DO NOT CHANGE THIS
\usepackage{aaai2026}  % DO NOT CHANGE THIS
\usepackage{times}  % DO NOT CHANGE THIS
\usepackage{helvet}  % DO NOT CHANGE THIS
\usepackage{courier}  % DO NOT CHANGE THIS
\usepackage[hyphens]{url}  % DO NOT CHANGE THIS
\usepackage{graphicx} % DO NOT CHANGE THIS
\urlstyle{rm} % DO NOT CHANGE THIS
\def\UrlFont{\rm}  % DO NOT CHANGE THIS
\usepackage{natbib}  % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
\usepackage{caption} % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
\frenchspacing  % DO NOT CHANGE THIS
\setlength{\pdfpagewidth}{8.5in}  % DO NOT CHANGE THIS
\setlength{\pdfpageheight}{11in}  % DO NOT CHANGE THIS
%
% These are recommended to typeset algorithms but not required. See the subsubsection on algorithms. Remove them if you don't have algorithms in your paper.
\usepackage{algorithm}
\usepackage{algorithmic}

%
% These are are recommended to typeset listings but not required. See the subsubsection on listing. Remove this block if you don't have listings in your paper.
\usepackage{newfloat}
\usepackage{listings}
\DeclareCaptionStyle{ruled}{labelfont=normalfont,labelsep=colon,strut=off} % DO NOT CHANGE THIS
\lstset{%
	basicstyle={\footnotesize\ttfamily},% footnotesize acceptable for monospace
	numbers=left,numberstyle=\footnotesize,xleftmargin=2em,% show line numbers, remove this entire line if you don't want the numbers.
	aboveskip=0pt,belowskip=0pt,%
	showstringspaces=false,tabsize=2,breaklines=true}
\floatstyle{ruled}
\newfloat{listing}{tb}{lst}{}
\floatname{listing}{Listing}
%
% Keep the \pdfinfo as shown here. There's no need
% for you to add the /Title and /Author tags.
\pdfinfo{
/TemplateVersion (2026.1)
}

% DISALLOWED PACKAGES
% \usepackage{authblk} -- This package is specifically forbidden
% \usepackage{balance} -- This package is specifically forbidden
% \usepackage{color (if used in text)
% \usepackage{CJK} -- This package is specifically forbidden
% \usepackage{float} -- This package is specifically forbidden
% \usepackage{flushend} -- This package is specifically forbidden
% \usepackage{fontenc} -- This package is specifically forbidden
% \usepackage{fullpage} -- This package is specifically forbidden
% \usepackage{geometry} -- This package is specifically forbidden
% \usepackage{grffile} -- This package is specifically forbidden
% \usepackage{hyperref} -- This package is specifically forbidden
% \usepackage{navigator} -- This package is specifically forbidden
% (or any other package that embeds links such as navigator or hyperref)
% \indentfirst} -- This package is specifically forbidden
% \layout} -- This package is specifically forbidden
% \multicol} -- This package is specifically forbidden
% \nameref} -- This package is specifically forbidden
% \usepackage{savetrees} -- This package is specifically forbidden
% \usepackage{setspace} -- This package is specifically forbidden
% \usepackage{stfloats} -- This package is specifically forbidden
% \usepackage{tabu} -- This package is specifically forbidden
% \usepackage{titlesec} -- This package is specifically forbidden
% \usepackage{tocbibind} -- This package is specifically forbidden
% \usepackage{ulem} -- This package is specifically forbidden
% \usepackage{wrapfig} -- This package is specifically forbidden
% DISALLOWED COMMANDS
% \nocopyright -- Your paper will not be published if you use this command
% \addtolength -- This command may not be used
% \balance -- This command may not be used
% \baselinestretch -- Your paper will not be published if you use this command
% \clearpage -- No page breaks of any kind may be used for the final version of your paper
% \columnsep -- This command may not be used
% \newpage -- No page breaks of any kind may be used for the final version of your paper
% \pagebreak -- No page breaks of any kind may be used for the final version of your paperr
% \pagestyle -- This command may not be used
% \tiny -- This is not an acceptable font size.
% \vspace{- -- No negative value may be used in proximity of a caption, figure, table, section, subsection, subsubsection, or reference
% \vskip{- -- No negative value may be used to alter spacing above or below a caption, figure, table, section, subsection, subsubsection, or reference

\setcounter{secnumdepth}{0} %May be changed to 1 or 2 if section numbers are desired.

\title{Detecting AI-Generated Text Using Fine-Tuned LLMs and Retrieval-Augmented Classification}

\author{
    Balasubramoni Sneha,
    Goel Aditya Kumar,
    Modi Twissa,\\
    Nargas Tanmay Sanjay,
    Singh Ananya,
    Yu Taek Lee
}
\affiliations{
    Nanyang Technological University\\
    BALA0055@e.ntu.edu.sg\\
    ADITYAKU003@e.ntu.edu.sg\\
    TWISSA001@e.ntu.edu.sg\\
    TANMAYSA001@e.ntu.edu.sg\\
    ANANYA029@e.ntu.edu.sg\\
    YUTAEK001@e.ntu.edu.sg
}

\begin{document}

\maketitle

\begin{abstract}
Large language models can now generate text that is nearly indistinguishable from human writing, and the proliferation of such content on social media platforms raises serious concerns about misinformation, opinion manipulation, and automated spam. Existing detection methods---including neural classifiers, zero-shot statistical tests, and watermarking---typically target a single generator and degrade when confronted with text from unseen models or adversarial perturbations. We propose a hybrid detection framework built around two complementary signals: QLoRA-fine-tuned large language model detectors (LLaMA 3.1-8B and Qwen 3-8B) and a retrieval-based KNN detector over a FAISS-indexed labeled corpus. We also study retrieval-augmented training variants in which each training example is paired with its nearest labeled neighbors, while the fine-tuned LLM itself receives only the raw post at inference time. We evaluate four model configurations and a retrieval-only baseline across three test scenarios: standard, cross-model, and cross-platform, reporting TPR@1\%FPR, AUROC, and Macro-F1. Our ablation study quantifies when retrieval improves generalization and shows that its benefit depends strongly on the underlying base model.
\end{abstract}

\section{Introduction}

Large language models (LLMs) such as GPT-4 \cite{openai2023gpt4} and Qwen \cite{qwen2024} can now generate text that is virtually indistinguishable from human writing across a wide range of domains and styles. Social media platforms---including Twitter, Reddit, and Telegram---are particularly vulnerable to misuse of these capabilities due to their high posting volume, low editorial oversight, and broad societal reach. The proliferation of AI-generated content on these platforms raises serious concerns about misinformation campaigns, opinion manipulation, and automated spam \cite{goldstein2023}, making reliable detection of machine-generated text an increasingly urgent problem.

A variety of detection approaches have been proposed. Statistical methods such as GLTR \cite{gehrmann2019gltr} analyze token-level likelihood under a reference language model. Neural classifiers, exemplified by the RoBERTa-based GPT-2 detector \cite{solaiman2019release}, learn discriminative features from labeled data. Zero-shot methods like DetectGPT \cite{mitchell2023detectgpt} avoid generator-specific training by exploiting probability curvature properties. Watermarking techniques \cite{kirchenbauer2023watermark} embed detectable signals during generation. However, a recurring limitation is that these methods are typically developed and evaluated against a single generator, and their performance degrades substantially when confronted with text from unseen AI models or adversarial perturbations \cite{dugan2024raid, sadasivan2023}.

To address this generalization challenge, we propose a hybrid detection framework that combines two complementary strategies. First, we fine-tune large language models (LLaMA 3.1-8B and Qwen 3-8B) using QLoRA \cite{dettmers2023qlora, hu2022lora} for binary classification of AI-generated versus human-written text. Second, we build a retrieval pipeline over a FAISS-indexed labeled corpus, which serves two roles in our system: a standalone KNN detector based on nearest-neighbor voting, and a source of retrieved examples for retrieval-augmented training variants. For the retrieval-augmented LLM variants, each training example is paired with its nearest labeled neighbors during fine-tuning, but at inference time the fine-tuned LLM receives only the raw post. This design lets us compare retrieval-based voting against model-based reasoning directly, while still testing whether retrieval-augmented training improves generalization to unseen generators and platforms.

We make the following contributions:
\begin{itemize}
    \item We propose a hybrid retrieval-augmented fine-tuning framework that combines QLoRA-adapted LLMs with embedding-based retrieval for AI-generated text detection, leveraging the complementary strengths of discriminative fine-tuning and adaptive retrieval.
    \item We conduct a comprehensive evaluation across three test scenarios---standard, cross-model, and cross-platform---using three datasets: MultiSocial \cite{multisocial2025}, HC3 \cite{guo2023hc3}, and an Adversarial QA corpus reserved for robustness evaluation.
    \item We present a systematic ablation study comparing fine-tuning-only, retrieval-only, and combined configurations across multiple base models, isolating the individual and combined contributions of each component as measured by TPR@1\%FPR, AUROC, and Macro-F1.
\end{itemize}

The remainder of this paper is organized as follows. The next section surveys related work on AI text detection, retrieval-augmented methods, and robustness evaluation. The Approach section describes our method, including the retrieval pipeline and fine-tuning procedure. The Experiments section presents the experimental setup, datasets, and results. The Analysis section provides discussion, and the final section concludes.

\section{Related Work}

\paragraph{AI-Generated Text Detection.}
Detecting machine-generated text has been studied through several complementary paradigms. Statistical approaches such as GLTR \cite{gehrmann2019gltr} visualize per-token likelihood under a language model to highlight unlikely word choices, providing an interpretable but coarse detection signal. Neural classifiers, most notably the RoBERTa-based detector released alongside GPT-2 \cite{solaiman2019release}, achieve strong accuracy on the generator they were trained on but degrade substantially when applied to text from newer or different models. Zero-shot methods address this limitation by avoiding generator-specific training: DetectGPT \cite{mitchell2023detectgpt} exploits the observation that model-generated text tends to lie in regions of negative log-probability curvature, and Fast-DetectGPT \cite{bao2024fastdetectgpt} reduces the computational cost of this approach via conditional probability estimation. Watermarking techniques \cite{kirchenbauer2023watermark} embed statistical signals during generation that can later be detected, but require cooperation from model providers and cannot be applied retroactively to existing content. A persistent limitation across these methods is that they are typically trained or calibrated on a single generator and evaluated within the same distribution, leaving cross-model generalization largely unaddressed. Our work tackles this gap by augmenting a fine-tuned classifier with retrieval-based evidence drawn from a multi-generator corpus.

\paragraph{Retrieval-Augmented Methods.}
Retrieval-augmented approaches to NLP were popularized by \citet{lewis2020rag}, who combined a dense retriever with a sequence-to-sequence generator for knowledge-intensive tasks. Subsequent work has extended retrieval augmentation to language model pre-training \cite{guu2020realm} and has shown that nearest-neighbor retrieval over a datastore can improve language modeling without additional training \cite{khandelwal2020knnlm}. Efficient similarity search at the scale of millions of vectors is enabled by libraries such as FAISS \cite{johnson2019faiss}. While most retrieval-augmented methods focus on improving generation quality or perplexity, their application to classification tasks remains underexplored. In our framework, we apply retrieval augmentation to binary classification: retrieved labeled examples serve as evidence for the classifier rather than as seeds for text generation, providing both an interpretability benefit and a mechanism for adapting to new generators by updating the retrieval database alone.

\paragraph{Social Media Text and Cross-Platform Detection.}
Social media text presents distinct challenges for AI-generated text detection: posts are typically short, informal, multilingual, and shaped by platform-specific conventions such as hashtags, mentions, and character limits. Prior work on social bot detection \cite{cresci2020} has focused on account-level behavioral signals rather than text-level linguistic analysis. Concerns about AI-driven influence operations on social platforms have been raised by \citet{goldstein2023}, yet most detection research evaluates on long-form, formal domains such as news articles or academic writing. The MultiSocial benchmark \cite{multisocial2025} addresses this gap by providing labeled AI-generated and human-written texts across three social media platforms (Twitter, Telegram, and Discord), seven AI models, and more than 24 languages. We build on this resource and additionally incorporate the HC3 corpus \cite{guo2023hc3} to evaluate cross-platform transfer, an axis that remains understudied in the detection literature.

\paragraph{Robustness and Adversarial Evaluation.}
The robustness of AI text detectors has come under scrutiny as adversarial strategies have proven effective at evading them. \citet{sadasivan2023} argue that reliable detection may become fundamentally difficult as language models improve, while \citet{krishna2023paraphrase} demonstrate that simple paraphrasing attacks can defeat most detectors---but also show that retrieval-based methods offer an effective defense, a finding that directly motivates our approach. The RAID benchmark \cite{dugan2024raid} formalizes adversarial evaluation with 11 generators and 12 attack types, revealing that many detectors lose significant accuracy under attack. Additionally, \citet{liang2023bias} highlight that existing detectors exhibit bias against non-native English writers, raising fairness concerns for deployment. In our experiments, we compare detector behavior across three settings---standard, cross-model, and cross-platform---to assess generalization beyond the training distribution.

\section{Approach}

This section describes the five detection configurations evaluated in this work.

\subsection{LLaMA}

This subsection describes the fine-tuned LLaMA classifier without retrieval augmentation.

\subsection{LLaMA + Retrieval-Augmented Classification}

This subsection describes the fine-tuned LLaMA classifier with retrieved labeled examples as additional evidence.

\subsection{Qwen}

This subsection describes the fine-tuned Qwen classifier without retrieval augmentation.

\subsection{Qwen + Retrieval-Augmented Classification}

This subsection describes the fine-tuned Qwen classifier with retrieved labeled examples as additional evidence.

\subsection{Simple KNN Voting}

This subsection describes the retrieval-only baseline that predicts labels by voting among nearest neighbors in the embedding space.

\section{Experiments}

This section presents the datasets, evaluation protocol, model settings, and experimental results.

\subsection{Data}

We use three datasets with distinct roles. MultiSocial \cite{multisocial2025} is the primary training corpus: a multilingual social-media collection of more than 472{,}000 labeled texts spanning three platforms (Twitter, Telegram, and Discord) and more than 24 languages. HC3 \cite{guo2023hc3} provides supplementary question-answer style data that broadens the writing styles seen during training. An Adversarial QA corpus is reserved strictly for evaluation and is not used during training. We unify these resources through a six-step preprocessing pipeline covering format unification, text normalization, label unification, embedding and retrieval preparation, retrieval-aware example construction, and final evaluation split generation. Labels are collapsed to a binary \textit{ai}/\textit{human} target, and the raw class distribution is highly imbalanced at approximately 87\% AI-generated versus 12.3\% human-written text.

To address this imbalance, we apply random undersampling to the AI class in the training data using a fixed seed, rebuild the FAISS \cite{johnson2019faiss} index on the balanced subset, and regenerate both retrieval-augmented and prompt-only training files. The resulting balanced training corpus contains 133{,}926 examples split evenly across AI and human labels. Validation and test sets retain the natural skew to better reflect real-world conditions. Embeddings are produced with Google's Gemini embedding-2 model at 768 dimensions, and the retrieval index is built from the training split only to prevent leakage. For retrieval-augmented training variants, each example is paired with its top-5 retrieved labeled neighbors; at inference time, however, the fine-tuned LLM path consumes only the raw post, while the retrieval-only KNN path queries the index directly. The preprocessing pipeline produces five evaluation splits---standard, cross-model, cross-platform, adversarial, and short-text---and we report results on the three core splits (standard, cross-model, and cross-platform) in this paper.

\subsection{Evaluation Method}

This subsection defines the test scenarios and metrics used to compare detector performance.

\subsection{Model Settings}

This subsection outlines the training and inference settings for the LLaMA, Qwen, and retrieval-based models.

\subsection{Results}

This subsection reports the performance of all evaluated configurations across the three test scenarios.

\section{Analysis}

This section analyzes performance trends, failure cases, and the contribution of retrieval across settings.

\section{Conclusion}

This section summarizes the main findings and discusses implications for robust AI-generated text detection.

\bibliography{aaai2026}

\end{document}
